---
title: "Causal Survival Analysis"
author: "Robin J. Evans"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Causal_Survival_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

We begin by loading the library.

```{r setup}
library(survivl)
```

## Time-Varying Covariates

If the treatment and covariates are allowed to be time-varying, then an ordinary
Cox model will not have any causal interpretation.  Instead we must use something
like a Cox _Marginal Structural Model_ (MSM) instead.  This assumes that the 
set of observed covariates is sufficient to control for confounding at any 
given timepoint, and that the causal quantity of interest is the marginal effect
of the treatment only on the outcome.

We can simulate from such a model using a copula for the outcome and 
covariates.  We select a normal covariate $Z$, a binary treatment $X$, and we 
must have an exponentially distributed outcome $Y$.  We also use a Gaussian
copula.

```{r MSM_forms}
formulas <- list(list(),
                 Z ~ X_l1,
                 X ~ Z_l0,
                 Y ~ X_l0,
                 cop ~ 1)
family <- list(integer(0), 1, 5, 3, 1)
```

The parameters to be specified depends upon the particular families used.
```{r MSM_pars}
pars <- list(Z = list(beta=c(1,0.5), phi=1),
             X = list(beta=c(-0.25,0.5)),
             Y = list(beta=c(1.1,-0.2), lambda0=0.01),
             cop = list(beta=1))
```

Now we call the function `MSM_samp`, which uses rejection sampling to obtain 
data from this model.
```{r MSM_datasamp}
set.seed(123)
dat <- coxSamp(1e4, T=20,
               formulas = formulas, 
               family = family, 
               pars = pars,
               link = list(character(0), "identity", "logit", "inverse"))
datl <- surv_to_long(dat)
```

## Fitting the Model

Since this is a time-varying model, we must use inverse probability weighting 
to obtain the parameters.  We recommend using the `survey` library's `svyglm`
function for inference.

```{r survey_load, message=FALSE}
library(survey)
library(dplyr)
```
Now, we fit a GLM to obtain the weighting to use with $X$.
```{r weights}
modX <- glm(X ~ Z, family=binomial, data=datl)
summary(modX)
pred <- predict(modX, type="response")
datl <- datl %>% mutate(datl, wts = X/pred + (1-X)/(1-pred))
```

At this point we can fit the model to obtain the coefficients for the effect
of $X$ on $Y$.
```{r MSM_fit}
mod <- svyglm(Y ~ X, 
              family=binomial(link=log), 
              design = svydesign(ids=datl$id,
                                 weights=datl$wts,
                                 data=datl))
summary(mod)
```

## A Naïve Approach

If we ignore the fact that the covariates are time-varying, we will obtain a biased estimate of the coefficient of $X$ on $Y$.

```{r naive_fit}
mod_naive <- svyglm(Y ~ X, 
              family=binomial(link=log), 
              design = svydesign(ids=datl$id,
                                 data=datl))
summary(mod_naive)
```
