---
title: "Causal Survival Analysis"
author: "Robin J. Evans"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Causal_Survival_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

We begin by loading the library.

```{r setup}
library(survivl)
```

## Time-Varying Covariates

If the treatment and covariates are allowed to be time-varying, then an ordinary
Cox proportional hazards model will not have any causal interpretation.  Instead 
we must use something like a Cox _Marginal Structural Model_ (MSM) instead.  This assumes that the 
set of observed covariates is sufficient to control for confounding at any 
given timepoint, and that the causal quantity of interest is the marginal effect
of the treatment only on the outcome.

We can simulate from such a model using a copula for the outcome and 
covariates.  We select a normal covariate $Z$, a binary treatment $X$, and we 
must have an exponentially distributed outcome $Y$.  We also use a Gaussian
copula.

```{r MSM_forms}
formulas <- list(list(),
                 Z ~ X_l1,
                 X ~ Z_l0,
                 Y ~ X_l0,
                 cop ~ 1)
family <- list(integer(0), 1, 5, 3, 1)
```

The parameters to be specified will depend upon the particular families used.  Note
that all the following distributions are conditional on $Y_{t-1} = 0$.  We
take 
\begin{align*}
Z_t \mid \overline{Z}_{t-1}, \overline{X}_{t-1}  &\sim N(1 + X_{t-1}/2, \, 1)\\
X_t \mid \overline{Z}_{t}, \overline{X}_{t-1}  &\sim \operatorname{Bern}(-1 + Z_t)\\
Y_t \mid do(\overline{X}_{t}) &= \mathbb{I}\{Y^*_t < 0\} \mid do(\overline{X}_{t})\\
Y^*_t \mid do(\overline{X}_{t}) &\sim \operatorname{Exp}(\exp\{1.1-X_t/2\}),
\end{align*}
and use a Gaussian copula with correlation $2\operatorname{expit}(1) - 1 = 0.462$ between 
$Z_t$ and $Y_t^*$.  
```{r MSM_pars}
pars <- list(Z = list(beta=c(1,0.5), phi=1),
             X = list(beta=c(-1,1)),
             Y = list(beta=c(1.1,-0.5), lambda0=1),
             cop = list(beta=1))
```

Now we call the function `msm_samp`, which obtains samples from this model.
```{r MSM_datasamp}
set.seed(123)
dat <- msm_samp(1e4, T=5,
                formulas = formulas, 
                family = family, 
                pars = pars,
                link = list(character(0), "identity", "logit", "inverse"))
datl <- surv_to_long(dat)
```

## Fitting the Model

Since this is a time-varying model, the obvious inference method is to use 
inverse probability weighting.  We recommend using the `survey` library's `svyglm`
function, since its standard errors account for the re-weighting.

```{r survey_load, message=FALSE}
library(survey)
library(dplyr)
options(digits=5)
```
Now, we fit a GLM to obtain the weighting to use with $X$.
```{r weights}
modX <- glm(X ~ Z, family=binomial, data=datl)
summary(modX)$coefficients
pred <- predict(modX, type="response")
datl <- datl %>% mutate(datl, wts = X/pred + (1-X)/(1-pred))
```

At this point we can fit the model to obtain the coefficients for the effect
of $X$ on $Y$.
```{r MSM_fit}
mod <- svyglm(I(1-Y) ~ X, 
              family=binomial(link=log), 
              design = svydesign(ids=datl$id,
                                 weights=datl$wts,
                                 data=datl))
summary(mod)$coefficients
```

## Naïve Approaches

If we ignore the fact that the covariates are time-varying, we will obtain a biased estimate of the coefficient of $X$ on $Y$.

```{r naive_fit, warning=FALSE}
mod_naive <- svyglm(I(1-Y) ~ X, 
              family=binomial(link=log), 
              design = svydesign(ids=datl$id,
                                 weights=~1,
                                 data=datl))
summary(mod_naive)$coefficients
```

Alternatively we can fit a conditional model.  Again, the estimate we obtain
is biased.

```{r cond_fit, warning=FALSE}
start <- mod_naive$coefficients
start <- c(start[1], 0, start[2])
mod_cond <- svyglm(I(1-Y) ~ Z + X, 
                   start = start,
                   family = binomial(link=log), 
                   design = svydesign(ids=datl$id,
                                      weights=~1,
                                      data=datl))
summary(mod_cond)$coefficients
```

