---
title: "Censoring"
author: "Robin J. Evans"
date: "31/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Censoring

```{r setup}
library(survivl)
```

The censoring mechanism is specified as another outcome variable.

```{r MSM_forms}
formulas <- list(list(),
                 Z ~ X_l1,
                 X ~ Z_l0,
                 list(Y ~ X_l0, C ~ X_l0),
                 cop ~ 1)
family <- list(integer(0), 1, 5, list(3,3), 1)
```

The parameters to be specified depends upon the particular families used.
```{r MSM_pars}
pars <- list(Z = list(beta=c(1,0.5), phi=1),
             X = list(beta=c(-0.25,0.5)),
             Y = list(beta=c(1.1,-0.2), lambda0=0.01),
             C = list(beta=c(1.1,-0.2), lambda0=0.01),
             cop = list(beta=1))
```

Now we call the function `MSM_samp`, which uses rejection sampling to obtain 
data from this model.
```{r MSM_datasamp}
set.seed(123)
dat <- coxSamp(1e3, T=5,
               formulas = formulas, 
               family = family, 
               pars = pars,
               link = list(character(0), "identity", "logit", c("inverse", "inverse")))
datl <- surv_to_long(dat)
```

## Fitting the Model

Since this is a time-varying model, we must use inverse probability weighting 
to obtain the parameters.  We recommend using the `survey` library's `svyglm`
function for inference.

```{r survey_load, message=FALSE}
library(survey)
library(dplyr)
```
Now, we fit a GLM to obtain the weighting to use with $X$.
```{r weights}
modX <- glm(X ~ Z, family=binomial, data=datl)
summary(modX)
pred <- predict(modX, type="response")
datl <- datl %>% mutate(datl, wts = X/pred + (1-X)/(1-pred))
```
