---
title: "Censoring"
author: "Robin J. Evans"
date: "31/05/2022"
output: html_document
vignette: >
  %\VignetteIndexEntry{Censoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Censoring

```{r load}
library(survivl)
```

The censoring mechanism is specified either as another outcome variable, or by
using the `Surv()` function in the outcome formula.

```{r MSM_forms}
formulas <- list(list(),
                 Z ~ X_l1,
                 X ~ Z_l0,
                 list(Y ~ X_l0, Cen ~ X_l0),
                 cop ~ 1)
family <- list(integer(0), 1, 5, list("weibull", "weibull"), 1)
```
The equivalent formulation using `Surv()` would replace the list in the fourth 
entry of `formulas` with `Surv(Y,Cen) ~ X_l0`.

The parameters to be specified depends upon the particular families used.
```{r MSM_pars}
pars <- list(Z = list(beta=c(-0.25,0.5), phi=1),
             X = list(beta=c(-0.5,0.5)),
             Y = list(beta=c(1.1,-0.2), shape=1, scale=0.5),
             Cen = list(beta=c(0.1,0), shape=1, scale=0.5),
             cop = list(Cen=list(Z=list(beta=0)), Y=list(Z=list(beta=0), Cen=list(beta=0))))
```

Now we call the function `msm_samp`, which uses inversion sampling to obtain 
data from a marginal structural model with the specified parameters.
```{r MSM_datasamp}
set.seed(123)
dat <- msm_samp(1e5, T=5,
                formulas = formulas, 
                family = family, 
                pars = pars)
# ,
#                 link = list(character(0), "identity", "logit", c("inverse", "inverse")))
datl <- surv_to_long(dat)
```

## Fitting the Model

Since this is a time-varying model, we use the appropriate formulation of the 
model in the `coxph()` function.  
```{r coxph0}
library(survival)
mod <- coxph(Surv(t, t_stop, Y) ~ X, data = datl, cluster = id)
summary(mod)
```

