---
title: "Complicated Copulas"
author: "Chase Mathis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Complicated Copulas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  warning = FALSE
)
```

```{r load, message=FALSE}
library(survivl)
library(dplyr)
library(ipw)
library(broom)
library(knitr)
library(dplyr)
library(survival)
```

This vignette will discuss the different ways a user can specify the dependence structure
between the outcome variable and covariates with increasing copmlexity. 

## A Simple Example
This example mimics the example in `Examples.Rmd`. We specify the copula as 
the last argument in the formulas list as a list itself. We specify that $Y$, the outcome,
has a copula with our time-varying covariate $Z$ only. The last entry in the family
variable specify that this is a Gaussian copula and the last entry in the parameters
list gives the beta value for the linear value.

Mathematically, for $j,k \in \{0,1,2,3,4,5,6\}^2$ the dependence between 
$$Y_j, Z_k$$ is the same and it is a gaussian copula with parameter $rho = 0.4$. 

```{r sim_data1}
formulas <- list(C ~ 1,
                 Z ~ X_l1 + C,
                 X ~ Z_l0 + C,
                 Y ~ X_l0 + C,
                 list(Y = list(Z ~1)))
family <- list(5,1,5,3,1)
link <- list("logit", "identity", "logit", "inverse")
pars <- list(C = list(beta=0),
             Z = list(beta = c(-1/2,1/2,0.25), phi=0.5),
             X = list(beta = c(0,1/2,1/10)),
             Y = list(beta = c(0.25,0.5,0.05), phi=1),
             cop = list(Y = list(Z = list(beta=0.8472979))))  # gives correlation 0.4

set.seed(123)
n <- 2e4
surv_model <- survivl_model(T=7, formulas=formulas, family=family, pars=pars, link=link)
dat <- rmsm(n, surv_model)
datl1 <- surv_to_long(dat)

head(cor(datl1 %>% select(T, Z)))
```

## Conditional Copulas
Similar to `causl`, we can set the dependence between $Y$ and $X$ to be conditional 
on some other variables. Here, we set the dependence between $Y_k$ and $Z_j$
to be a student's t copula conditional on $C, X_0$. Our latent correlation
variable $\eta$ is defined as $\eta_i = 2 * \rm{expit}(0*1 + 0.5*C_i + 0.8472979 * X_{0,i}) - 1$ and for each observation of $Y_k, Z_j$ the correlation parameter is $\eta_i$.
```{r sim_data2}
formulas <- list(C ~ 1,
                 Z ~ X_l1 + C,
                 X ~ Z_l0 + C,
                 Y ~ X_l0 + C,
                 list(Y = list(Z ~ C + X_0)))
family <- list(5,1,5,3,1)
link <- list("logit", "identity", "logit", "inverse")
pars <- list(C = list(beta=0),
             Z = list(beta = c(-1/2,1/2,0.25), phi=0.5),
             X = list(beta = c(0,1/2,1/10)),
             Y = list(beta = c(0.25,0.5,0.05), phi=1),
             cop = list(Y = list(Z = list(beta=c(0, 0.5, 0.8472979), df = 5))))  

set.seed(123)
n <- 2e4
surv_model <- survivl_model(T=7, formulas=formulas, family=family, pars=pars, link=link)
dat <- rmsm(n, surv_model)
datl2 <- surv_to_long(dat)

head(cor(datl2 %>% select(T, Z)))
```

## Multiple Time Varying Covariates

This example includes multiple time-varying confounders. We have both $Z1$ and $Z2$ and therefore
need to model the dependence between $Y$ and $Z1$ and between $Y$ and $Z2 \mid Z1$.
The ordering of this is natural from the order of the time-varying covariates.

Previously, we had the same copula for all time steps $t$, but now we are specifying
a markov formula for the copula like the other formulas. So, the dependency 
between $Y_k$ and $Z2_j$ is conditioned on $Z1_j$. 

```{r sim_data3}
n <- 2e4
formulas <- list(C ~ 1,
                 list(Z1 ~ X_l1 + C,Z2 ~ X_l1 + Z1_l0),
                 X ~ Z1_l0 + Z2_l0 + C,
                 Y ~ X_l0 + C,
                 list(Y = list(Z1 ~1, Z2 ~ Z1_l0)))
family <- list(5,c(5,1),5,3,c(1,1))
link <- list("logit", c("logit","identity"), "logit", "inverse")
pars <- list(C = list(beta=0),
             Z1 = list(beta = c(-1/2,1/2,0.25), phi=0.5),
             Z2 = list(beta = c(0, -1/4, 2/5), phi = 0.65),
             X = list(beta = c(0,1/4, 1/2,1/10)),
             Y = list(beta = c(0.25,0.5,0.05), phi=1),
             cop = list(Y = list(Z1 = list(beta= 0.847), Z2 = list(beta=c(0.5,0.847))))) 

set.seed(123)
surv_model <- survivl_model(T=7, formulas=formulas, family=family, pars=pars, link=link)
dat <- rmsm(n, surv_model)
datl3 <- surv_to_long(dat)

cor(datl3 %>% select(T, Z1, Z2))

```
## Conditioning on Treatments
In addition to conditioning on other time-varying covariates,
the user can also condition on past treatments. This example extends on our previous,
but now the dependence between $Z1_j$ and $Y_k$ depends on $X_{j-1}$ and the dependence
between $Y_k$ and $Z2_j$ depends on $Z1_j$ (like before) and on $X_0$. This has a lot
of moving parts and a probably over-complicated model, but its meant to illustrate the flexiblity
of the copula construction.

 


```{r sim_data4}
n <- 2e4
formulas <- list(C ~ 1,
                 list(Z1 ~ X_l1 + C,Z2 ~ X_l1 + Z1_l0),
                 X ~ Z1_l0 + Z2_l0 + C,
                 Y ~ X_l0 + C,
                 list(Y = list(Z1 ~X_l1, Z2 ~ Z1_l0 + X_0)))
family <- list(5,c(5,1),5,3,c(1,1))
link <- list("logit", c("logit","identity"), "logit", "inverse")
pars <- list(C = list(beta=0),
             Z1 = list(beta = c(-1/2,1/2,0.25), phi=0.5),
             Z2 = list(beta = c(0, -1/4, 2/5), phi = 0.65),
             X = list(beta = c(0,1/4, 1/2,1/10)),
             Y = list(beta = c(0.25,0.5,0.05), phi=1),
             cop = list(Y = list(Z1 = list(beta= c(0,0.847)), 
                                           Z2 = list(beta=c(0, 0.5,0.847)))))

set.seed(123)
surv_model <- survivl_model(T=7, formulas=formulas, family=family, pars=pars, link=link)
dat <- rmsm(n, surv_model)
datl4 <- surv_to_long(dat)

cor(datl4 %>% select(T, Z1, Z2))
```

## Verifying The Correct Marginal Quantity
We can estimate the marginal causal quantity $p(y \mid \rm{do}(X))$ with IPW. We
calculate the models and print the estimates at the end. All estimates are within 2
s.es.
```{r}

# Compute IPW weights
ipw_weights <- ipwtm(
  exposure = X,
  family = "binomial", # all of our treatments are binary
  link = "logit",
  numerator = ~C,
  denominator = ~Z,
  id = id,
  timevar = t,
  type = "all",
  data = datl1
)
datl1$wt <- ipw_weights$ipw.weights
sumY1 <- coxph(Surv(t, t_stop, Y) ~ X + C, data = datl1, timefix = FALSE, id =id,
              weight = wt)


# Compute IPW weights
ipw_weights <- ipwtm(
  exposure = X,
  family = "binomial", # all of our treatments are binary
  link = "logit",
  numerator = ~C,
  denominator = ~Z,
  id = id,
  timevar = t,
  type = "all",
  data = datl2
)
datl2$wt <- ipw_weights$ipw.weights
sumY2 <- coxph(Surv(t, t_stop, Y) ~ X + C, data = datl2, timefix = FALSE, id =id,
              weight = wt)
```

```{r}
# Compute IPW weights
ipw_weights <- ipwtm(
  exposure = X,
  family = "binomial", # all of our treatments are binary
  link = "logit",
  numerator = ~C,
  denominator = ~Z1 + Z2,
  id = id,
  timevar = t,
  type = "all",
  data = datl3
)
datl3$wt <- ipw_weights$ipw.weights
sumY3 <- coxph(Surv(t, t_stop, Y) ~ X + C, data = datl3, timefix = FALSE, id =id,
              weight = wt)

# Compute IPW weights
ipw_weights <- ipwtm(
  exposure = X,
  family = "binomial", # all of our treatments are binary
  link = "logit",
  numerator = ~C,
  denominator = ~Z1 + Z2,
  id = id,
  timevar = t,
  type = "all",
  data = datl4
)
datl4$wt <- ipw_weights$ipw.weights
sumY4 <- coxph(Surv(t, t_stop, Y) ~ X + C, data = datl4, timefix = FALSE, id =id,
              weight = wt)

```

```{r pretty-print-output, echo=FALSE}

# Tidy the model summaries
sumY_list <- list(sumY1, sumY2, sumY3, sumY4)
model_names <- paste0("Model ", 1:4)

tidy_list <- lapply(sumY_list, function(fit) {
  broom::tidy(fit) %>%
    select(term, estimate, std.error, statistic, p.value)
})

# Add model name to each
for (i in seq_along(tidy_list)) {
  tidy_list[[i]]$Model <- model_names[i]
}

# Combine into one data frame
results_tbl <- bind_rows(tidy_list) %>%
  select(Model, everything())

# Print as a table
kable(results_tbl, digits = 3, caption = "Cox Model Results for sumY1 - sumY4")
```