---
title: "Examples"
author: "Robin J. Evans"
date: "24/05/2022"
output: html_document
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

```{r load, message=FALSE}
library(survivl)
library(dplyr)
library(ipw)
```

## Two Examples

```{r params1}
formulas <- list(C ~ 1,
                 Z ~ X_l1 + C,
                 X ~ Z_l0 + C,
                 Y ~ X_l0 + C,
                 cop ~ 1)
family <- list(5,1,5,3,1)
link <- list("logit", "identity", "logit", "inverse")
pars <- list(C = list(beta=0),
             Z = list(beta = c(-1/2,1/2,0.25), phi=0.5),
             X = list(beta = c(0,1/2,1/10)),
             Y = list(beta = c(0.05,0.5,0.05), phi=1),
             cop = list(beta=0.8472979))  # gives correlation 0.4
```

## Simulating Data

```{r sim_dat1}
set.seed(123)
n <- 1e4
surv_model <- survivl_model(T=7, formulas=formulas, family=family, pars=pars, link=link)
dat <- rmsm(n, surv_model)
datl <- surv_to_long(dat)
```

```{r plot_dat1, message=FALSE}
library(ggfortify)
library(survival)
survplot <- survfit(Surv(t, t_stop, Y) ~ C, data=datl)
autoplot(survplot)
```

We recommend using `ipw::ipwtm` function to calculate ipw weights for time varying confounding. 
```{r glms1}
temp <- ipwtm(
  exposure = X,
  family = "binomial",
  numerator = ~ 1,
  denominator = ~ Z,
  link = "logit",
  id = id,
  timevar = t,
  type = "all",
  data = datl)


datl$wt <- temp$ipw.weights
mod_correct <- coxph(Surv(t, t_stop,Y)~X + C, id = id, data = datl,
                               weights = datl$wt, timefix = FALSE)

mod_naive <- coxph(Surv(t, t_stop,Y)~ X + C, id = id, data = datl,
                               weights = rep(1,nrow(datl)), timefix = FALSE)

```

```{r tabs, eval=FALSE, echo=FALSE}
tab_correct <- cbind(pars$Y$beta[-c(1)], -mod_correct$coef, summary(mod_correct)$coef[,-c(1,3)])
tab_correct[,4] <- pt(abs((tab_correct[,2]-tab_correct[,1])/tab_correct[,3]), df=n-4, lower.tail = FALSE)
tab_naive <- cbind(pars$Y$beta[-c(1)], -mod_naive$coef, summary(mod_naive)$coef[,-c(1,3)])
tab_naive[,4] <- pt(abs((tab_naive[,2]-tab_naive[,1])/tab_naive[,3]), df=n-4, lower.tail = FALSE)
library(kableExtra)
kableExtra::kbl(tab_naive, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
kableExtra::kbl(tab_correct, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
```

### Example 2 (Gamma distributed requires by hand ipw calculation)

We assume:
 - $C$ is a Bernoulli r.v. with parameter $1/2$;
 - $Z_t$ is a normal random variable with mean $C - 1/2 + X_{t-1}$;
 - $X_t$ is a Gamma random variable with log mean $Z_t/3 + C/3$;
 - $Y_t$ is an underlying exponential distribution with rate $X_t + C$;
 - the copula between $Y_t$ and $Z_t$ is Gaussian with fixed correlation.

```{r params2}
formulas <- list(C ~ 1,
                 Z ~ X_l1 + C,
                 X ~ Z_l0 + C,
                 Y ~ X_l0 + C,
                 cop ~ 1)
family <- list(5,1,3,3,1)
link <- list("logit", "identity", "log", "inverse")
pars <- list(C = list(beta=0),
             Z = list(beta = c(-1/2,1/4,1), phi=0.5),
             X = list(beta = c(0,1/4,1/20), phi=0.25),
             Y = list(beta = c(0,0.2,0.05), phi=1),
             cop = list(beta=1))
```

### Simulating Data

```{r sim_dat2}
set.seed(123)
surv_model2 <- survivl_model(T=7, formulas=formulas, family=family, pars=pars, link=link)
dat2 <- rmsm(1e4, surv_model2)
datl2 <- surv_to_long(dat2)
```

```{r plot_dat2, message=FALSE}
library(ggfortify)
library(survival)
survplot <- survfit(Surv(t, t_stop, Y) ~ C, data=datl2)
autoplot(survplot)
```

The `ipwtm` function allows for certain families, but not the gamma family. 
Since we assume our exposure `X` is Gamma (conditional on `Z, C`), we must calcualte
the ipw by hand.
```{r glms2}
mod_den <- glm(X ~ Z, family = Gamma(link = "log"), data = datl2)
mu_den  <- fitted(mod_den)
phi_den <- summary(mod_den)$dispersion

## 2. Numerator model
mod_num <- glm(X ~ 1,   family = Gamma(link = "log"), data = datl2)
mu_num  <- fitted(mod_num)
phi_num <- summary(mod_num)$dispersion

num_dens <- dgamma(datl2$X, shape = 1/phi_num, rate = 1/(phi_num*mu_num))
den_dens <- dgamma(datl2$X, shape = 1/phi_den, rate = 1/(phi_den*mu_den))

datl2 <- datl2 %>%
  mutate(ipw_instantaneous = num_dens / den_dens) %>%      # instantaneous weight
  group_by(id) %>%
  arrange(t) %>%
  mutate(ipw = cumprod(ipw_instantaneous)) %>% 
  ungroup()
```

Fit the coxph model with and without weights, then recover estimates.
```{r coxph2}
mod_correct <- coxph(Surv(t, t_stop,Y)~ X + C, id = id, data = datl2,
                               weights = ipw, timefix = FALSE)

mod_naive <- coxph(Surv(t, t_stop,Y)~ X + C, id = id, data = datl2,
                               weights = rep(1,nrow(datl2)), timefix = FALSE)

```

```{r tabs2, eval=FALSE, echo=FALSE}
tab_correct <- cbind(pars$Y$beta[-c(1)], -mod_correct$coef, summary(mod_correct)$coef[,-c(1,3)])
tab_correct[,4] <- pt(abs((tab_correct[,2]-tab_correct[,1])/tab_correct[,3]), df=n-4, lower.tail = FALSE)
tab_naive <- cbind(pars$Y$beta[-c(1)], -mod_naive$coef, summary(mod_naive)$coef[,-c(1,3)])
tab_naive[,4] <- pt(abs((tab_naive[,2]-tab_naive[,1])/tab_naive[,3]), df=n-4, lower.tail = FALSE)
library(kableExtra)
kableExtra::kbl(tab_naive, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
kableExtra::kbl(tab_correct, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
```

