---
title: "Marginal Structural Survival Models"
author: "Robin Evans"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Marginal Structural Models

In this vignette we consider _marginal structural models_ for time-to-event 
data.  Begin by loading the library.
```{r load}
library(survivl)
```

Then we define our formulas and family variables.  We will have one baseline
variable, and a single time-varying covariate, treatment and outcome.  More 
specifically, let:
\begin{align*}
W &\sim \operatorname{Gamma}(5/2, \, 2)  \\
Z_t \mid \overline{Z}_{t-1}, \overline{X}_{t-1} &\sim N(0.7 Z_{t-1} + 0.2 X_t, \, 1)\\
X_t \mid \overline{Z}_{t}, \overline{X}_{t-1} &\sim \operatorname{Bernoulli}(\operatorname{expit}(0.5 X_t + 0.25 Z_{t-1}))\\ 
P(Y_t = 1 \mid do(\overline{X}_{t}), W) &= \exp\{ - X_t + W/10\},
\end{align*}
and assume that there is a Gaussian pair-copula linking $Y_t$ with $Z_{t-1},Z_t$,
which also depends upon $W$.  We assume that the parameter for the $Y_t$-$Z_{t-1}$
interaction is 0.2, and that for $Y_t$-$Z_{t}$ given $Z_{t-1}$ is 0.3.

```{r formulas}
forms <- list(W ~ 1,
              Z ~ Z_l1 + X_l1,
              X ~ X_l1 + Z_l0,
              Y ~ W + X_l0,
              list(Y = list(Z ~ W)))
fams <- list(3, 1, 5, 3, c(1))
pars <- list(W = list(beta=log(5), phi=1/2),
             Z = list(beta=c(0,0.7,0.2), phi=1),
             X = list(beta=c(0,0.25,0.5)), 
             Y = list(beta=c(2,-1/10,1/5), phi=1),
             cop = list(Y=list(Z=list(beta=c(0.5,0.5))))
)
```
We can now simulate $n=10^3$ observations from the model over 10 time points.
```{r sim_data}
n <- 1e3
set.seed(123)
dat <- cox_samp(n, T = 10, formulas = forms, family = fams, pars = pars, 
                method = "inversion")
datl <- surv_to_long(dat, lag=1)
```
### Checking the fit

We can check that our simulation has the correct parametric distributions.  We
use a Kolmogorov-Smirnov test to check the distribution of $W$, and then 
regressions to test the distributions of the $Z$s and $X$s.

```{r check_fits}
ks.test(dat$W, "pgamma", scale=5/2, shape=2)
summary(glm(Z ~ W + Z_l1 + X_l1, data=datl))
summary(glm(X ~ W + X_l1 + Z, data=datl, family=binomial))
```
Indeed, each of these variables is consistent with the parameters we used.  For
the survival outcome, we can use inverse probability weighting.  The `survey` 
package can account for the additional uncertainty introduced by weighting.
```{r survey}
library(survey)
```
Now we estimate the propensity scores, and fit the IPW model.
```{r ipw}
ps <- predict(glm(X ~ W + X_l1 + Z, data=datl, family=binomial), type = "response")
wt <- datl$X/ps + (1-datl$X)/(1-ps)
svyglm(Y ~ W + X, design=svydesign(~1, data=datl, family=binomial(log), weights=wt))
```

# Multiple time-varying covariates

We may well be interested in simulating more than one time-varying covariate. We
can easily allow for this, first by simply specifying another formula in the second 
element of our list...

```{r mv_forms}
forms2 <- forms
forms2[[2]] <- list(Z1 ~ Z1_l1 + Z2_l1 + X_l1, 
                    Z2 ~ Z2_l1 + X_l1 + Z1_l0)
forms2[[3]] <- X ~ W + Z1_l0 + Z2_l0
fams2 <- fams
fams2[[2]] <- c(1,1)
pars2 <- pars
pars2 <- c(pars2,
           list(Z1 = list(beta=c(0,0.7,0.3,0.2), phi=1),
                Z2 = list(beta=c(0,0.5,0.2,0.1), phi=1)))
pars2$X <- list(beta = c(0,0.25,0.1,0.3))
```
... and then by modifying the elements relating to the copula.
```{r mf_cop}
forms2[[5]] <- list(Y = list(Z1 ~ W, Z2 ~ W))
fams2[[5]] <- c(1, 1)
pars2$cop <- list(Y=list(Z1=list(beta=c(0.5,0.2)),
                         Z2=list(beta=c(0.25,0))))
```
We can then simulate as before.
```{r sim_dat_mv}
dat2 <-  cox_samp(n, T = 10, formulas = forms2, family = fams2, pars = pars2, 
                method = "inversion")
dat2l <- surv_to_long(dat2, lag=1)
```
