---
title: "Marginal Structural Survival Models"
author: "Robin Evans"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{MSMs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Marginal Structural Models

In this vignette we consider _marginal structural models_ for time-to-event 
data.  Begin by loading the library.
```{r load}
library(survivl)
library(survival) # for coxph
options(digits=4)
```

Then we define our formulas and family variables.  We will have one baseline
variable, and a single time-varying covariate, treatment and outcome.  More 
specifically, let:
\begin{align*}
W &\sim \operatorname{Gamma}(1/2, \, 2)  \\
Z_t \mid \overline{Z}_{t-1}, \overline{X}_{t-1} &\sim N(0.7 Z_{t-1} + 0.2 X_{t-1}, \, 1)\\
X_t \mid \overline{Z}_{t}, \overline{X}_{t-1} &\sim \operatorname{Bernoulli}(\operatorname{expit}(0.5 X_{t-1} + 0.25 Z_t))\\ 
P(Y_t = 1 \mid do(\overline{X}_{t}), W) &= \exp\{ - X_t + W/10\},
\end{align*}
and assume that there is a Gaussian pair-copula linking $Y_t$ with $Z_{t-1},Z_t$,
which also depends upon $W$.  We assume that the parameter for the $Y_t$-$Z_{t-1}$
interaction is 0.2, and that for $Y_t$-$Z_{t}$ given $Z_{t-1}$ is 0.3.

```{r formulas}
forms <- list(W ~ 1,
              Z ~ Z_l1 + X_l1,
              X ~ X_l1 + Z_l0,
              Y ~ W + X_l0,
              list(Y = list(Z ~ W)))
fams <- list(3, 1, 5, 3, c(1))
pars <- list(W = list(beta=0, phi=1/2),
             Z = list(beta=c(0,0.7,0.2), phi=1),
             X = list(beta=c(-0.5,0.25,0.5)), 
             Y = list(beta=c(1,-1/10,1/5), phi=1),
             cop = list(Y=list(Z=list(beta=c(0.5, 0.0))))
)
link <- list("log", "identity", "logit", "inverse")
surv_model <- survivl_model(T = 10, formulas = forms, family = fams, pars = pars, link=link)
```
We can now simulate $n=10^5$ observations from the model over 10 time points.
```{r sim_data}
n <- 1e5
set.seed(124)
dat <- rmsm(n, surv_model)
datl <- surv_to_long(dat, lag=1)
```
### Checking the fit

We can check that our simulation has the correct parametric distributions.  We
use a Kolmogorov-Smirnov test to check the distribution of $W$, and then 
regressions to test the distributions of the $Z$s and $X$s.

```{r check_fits}
ks.test(dat$W, "pgamma", scale=1/2, shape=2)
summary(glm(Z ~ W + Z_l1 + X_l1, data=datl))$coefficients
summary(glm(X ~ W + X_l1 + Z, data=datl, family=binomial))$coefficients
```
Indeed, each of these variables is consistent with the parameters we used.  For
the survival outcome, we can use inverse probability weighting.  The `ipw` 
package can estimate ipw weights for time varying confounders like `Z`.

```{r}
temp <- ipw::ipwtm(
  X,
  family = "binomial",
  link = "logit",
  numerator = ~1,
  denominator = ~ Z,
  id = id,
  type = "all",
  timevar = t,
  data = datl,
)
datl$wt <- temp$ipw.weights
mod_correct <- coxph(Surv(t, t_stop, Y) ~ W + X, weights = wt, data = datl,
                     timefix = FALSE)
mod_naive <- coxph(Surv(t, t_stop, Y) ~ W + X, weights = rep(1, nrow(datl)), data = datl,
                   timefix = FALSE)
```

```{r}
tab_correct <- cbind(pars$Y$beta[-c(1)], -mod_correct$coef, summary(mod_correct)$coef[,-c(1,3)])
tab_correct[,4] <- pt(abs((tab_correct[,2]-tab_correct[,1])/tab_correct[,3]), df=n-4, lower.tail = FALSE)
tab_naive <- cbind(pars$Y$beta[-c(1)], -mod_naive$coef, summary(mod_naive)$coef[,-c(1,3)])
tab_naive[,4] <- pt(abs((tab_naive[,2]-tab_naive[,1])/tab_naive[,3]), df=n-4, lower.tail = FALSE)
library(kableExtra)
kableExtra::kbl(tab_naive, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
kableExtra::kbl(tab_correct, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
```

# Multiple time-varying covariates

We may well be interested in simulating more than one time-varying covariate. 
A new update allows us to update our `surv_model` object without redefining it.

```{r mv_forms}
forms2 <- forms
forms2[[2]] <- list(Z1 ~ Z1_l1 + Z2_l1 + X_l1, 
                    Z2 ~ Z2_l1 + X_l1 + Z1_l0)
forms2[[3]] <- X ~ W + Z1_l0 + Z2_l0
fams2 <- fams
fams2[[2]] <- c(1,1)
pars2 <- pars
pars2 <- c(pars2,
           list(Z1 = list(beta=c(0,0.7,0.3,0.2), phi=1),
                Z2 = list(beta=c(0,0.5,0.2,0.1), phi=1)))
pars2$X <- list(beta = c(0,0.25,0.1,0.3))
link[[2]] <- rep("identity", 2)
```
... and then by modifying the elements relating to the copula.
```{r mf_cop}
forms2[[5]] <- list(Y = list(Z1 ~ W, Z2 ~ W))
fams2[[5]] <- list(c(1, 1))
pars2$cop <- list(Y=list(Z1=list(beta=c(0.5,0.2)),
                         Z2=list(beta=c(0.25,0))))
# should be able to just use modify
surv_model2 <- modify.survivl_model(surv_model, formulas = forms2,
                                    family = fams2,
                                    pars = pars2, link = link)
```
We can then simulate as before.
```{r sim_dat_mv}
dat2 <-  rmsm(n, surv_model2)
dat2l <- surv_to_long(dat2, lag = 1)
```

Checking the distributions, these all seem fine.
```{r check_ZX_mv, eval=FALSE}
summary(glm(Z1 ~ W + Z1_l1 + Z2_l1 + X_l1, data=dat2l))$coef
summary(glm(Z2 ~ W + Z2_l1 + X_l1 + Z1, data=dat2l))$coef
summary(glm(X ~ W + Z1 + Z2, data=dat2l, family=binomial))$coef
```

For the outcome, we again apply IPW, and obtain:
```{r}
temp <- ipwtm(
  exposure = X,
  family = "binomial",
  link = "logit",
  numerator = ~W,
  denominator = ~ W + Z1 + Z2 + X_l1,
  id = id,
  timevar = t,
  type = "all",
  data = dat2l
)
dat2l$wt <- temp$ipw.weights

mod_correct <- coxph(Surv(t, t_stop, Y) ~ W + X,  weights = dat2l$wt,
                     data = dat2l, timefix = FALSE)
mod_naive <- coxph(Surv(t, t_stop, Y) ~ W + X, weights = rep(1, nrow(dat2l)),
                     data = dat2l, timefix = FALSE)
```

```{r check_XY_mv, eval=TRUE, warn=FALSE}
tab_correct <- cbind(pars$Y$beta[-c(1)], -mod_correct$coef, summary(mod_correct)$coef[,-c(1,3)])
tab_correct[,4] <- pt(abs((tab_correct[,2]-tab_correct[,1])/tab_correct[,3]), df=n-4, lower.tail = FALSE)
tab_naive <- cbind(pars$Y$beta[-c(1)], -mod_naive$coef, summary(mod_naive)$coef[,-c(1,3)])
tab_naive[,4] <- pt(abs((tab_naive[,2]-tab_naive[,1])/tab_naive[,3]), df=n-4, lower.tail = FALSE)
library(kableExtra)
kableExtra::kbl(tab_naive, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
kableExtra::kbl(tab_correct, digits = c(2,3,5,2), booktabs=TRUE, format="latex")  %>%  
  kableExtra::add_header_above(c("Coef","Truth","Est.", "Std. Err.", "p-value"))
```


