---
title: "Plasmode"
author: "Chase Mathis"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{MSMs}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Marginal Structural Models with Plasmode (Existing Data)

In this vignette, we consider *marginal structural models* for
time-to-event data under the plasmode setting. This means there exists
data already and we continue to simulate from that pre-existing data.
Begin by loading the library.

```{r load}
library(survivl)
library(ipw)
rho_to_beta <- function(rho){
  x <- (rho + 1)/2
  return(log(x/(1-x)))
}
options(digits=4)
```

Then we define our formulas and family variables. We will have two baseline variables $X_1$ and $X_2$, and two additional baseline covariates $B_1$ and $B_2$. More specifically, let:
\begin{align*}
X_1 &\sim N(0,1) \\
X_2 &\sim \text{Bernoulli}(0.5) \\
B_1 \mid X_2 &\sim N(0.4X_2 - 0.2, 1) \\
B_2 \mid X_1 &\sim N(0.2X_1, 1) \\
Z_{1,0} \mid X_1 &\sim N(0.2X_1, 1) \\
Z_{2,0} \mid X_2 &\sim \text{Bernoulli}(\text{expit}(-0.2 + 0.4X_2)) \\
A_0 \mid X_1,X_2,B_1,Z_{1,0},Z_{2,0} &\sim \text{Bernoulli}(\text{expit}(-1 + 0.1X_1 + 0.15X_2 + 0.1B_1 + 0.3Z_{1,0} + 0.3Z_{2,0}))
\end{align*}


We have 10,000 observations of existing data.
```{r existing data}
n <- 1e4
dat <- data.frame(X1=rnorm(n,mean = 0, sd = 1),
                  X2 = rbinom(n, size = 1, prob = 0.5))
dat$B1 <- rnorm(n, 0.4*dat$X2 -0.2, sd = 1)

dat[["Z1_0"]] <- rnorm(n, 0.2*dat$X1, sd = 1)
dat[["Z2_0"]] <- rbinom(n, 1, plogis(-0.2 + 0.4 * dat$X2))
dat[["A_0"]] = rbinom(n, 1, plogis(-1 + 0.1 * dat$X1 + 0.15*dat$X2 + 0.1*dat$B1 + 0.3*dat[["Z1_0"]] + 0.3* dat[["Z2_0"]]))

```

Then we define the time-varying covariates and outcome. Let:
\begin{align*}
Z_{1,t} \mid B_2, Z_{1,t-1}, A_{t-1} &\sim N(0.3 + 0.4B_2 + 0.7Z_{1,t-1} - 0.6A_{t-1}, 1) \\
Z_{2,t} \mid B_2, Z_{2,t-1}, A_{t-1} &\sim \text{Bernoulli}(\text{expit}(-0.2 + 0.4B_2 + Z_{2,t-1} - 0.6A_{t-1})) \\
A_t \mid X_1,X_2,B_1,Z_{1,t},Z_{2,t},A_{t-1} &\sim \text{Bernoulli}(\text{expit}(-1 + \gamma_1X_1 + \gamma_2X_2 + \gamma_3B_1 + \gamma_4Z_{1,t} + \gamma_5Z_{2,t} + A_{t-1})) \\
Y \mid A_0 &\sim \text{Exponential}(\text{rate}=1, \text{scale}=\exp(0.5 + 0.2A_0))
\end{align*}

The copula structure between $Y$ and $(Z_1,Z_2)$ is defined through a t-copula with 5 degrees of freedom and correlation parameters $\rho_{Z_1} = -0.6$ and $\rho_{Z_2} = 0.2$.

```{r formulas}
gamma1 <- 0.1
gamma2 <- 0.15
gamma3 <- 0.1
gamma4 <- 0.3
gamma5 <- 0.3

formulas <- list(list(B2 ~ X1),
                 list(Z1 ~  B2 + Z1_l1 + A_l1, Z2 ~ B2 + Z2_l1 + A_l1),
                 A ~ X1 + X2 + B1 + Z1_l0 + Z2_l0 + A_l1,
                 Y ~ A_l0,
                 list(Z1 ~ 1, Z2 ~ 1))

family <- list(list(1),
               list(1,5),
               5,
               3,
               list(2,2))
pars <- list(B2 = list(beta = c(-0.2, 0.4), phi = 1),
             Z1 = list(beta=c(0.3,0.4, 0.7,-0.6 ), phi=1),
             Z2 = list(beta = c(-0.2,0.4,1,-0.6)),
             A = list(beta=c(-1, gamma1, gamma2, gamma3, gamma4, gamma5, 1)),
             Y = list(beta = c(0.5,0.2), phi = 1), # negative 1/exp
             cop = list(Y = list(Z1 = list(beta= rho_to_beta(-0.6), par2 = 5),
                                 Z2 = list(beta = rho_to_beta(0.2), par2 = 5))))

surv_model <- survivl_model(formulas=formulas, family=family,
                    pars=pars, T = 5, dat = dat, qtls = NULL, # no quantiles given
                    link = list(list("identity"),
                    list("identity", "logit"),
                    "logit", "log"))
```

We simulate the rest of the data `dat2`, starting with `dat`.
```{r}
dat2 <- rmsm(n, surv_model)
```

We are able to recover the causal estimand. 
```{r}
datl <- surv_to_long(dat2)


temp <- ipwtm(
  exposure = A,
  family = "binomial",
  numerator = ~ 1,
  denominator = ~ Z1 +Z2,
  link = "logit",
  id = id,
  timevar = t,
  type = "all",
  data = datl)


datl$wt <- temp$ipw.weights
glmY <- suppressWarnings(coxph(Surv(t, t_stop,Y)~ A, id = id, data = datl,
                               weights = datl$wt, timefix = FALSE))
sumY <- summary(glmY)
sumY
```

