---
title: "Plasmode"
author: "Chase Mathis"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{MSMs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Marginal Structural Models

In this vignette, we consider _marginal structural models_ for time-to-event 
data under the plasmode setting. This means there exists
data already and we continue to simulate from that pre-existing data. Begin by loading the library.
```{r load}
library(survivl)
options(digits=4)
```

Then we define our formulas and family variables.  We will have one baseline
variable, and a single time-varying covariate, treatment and outcome.  More 
specifically, let:
\begin{align*}
W &\sim \operatorname{Gamma}(1/2, \, 2)  \\
Z_t \mid \overline{Z}_{t-1}, \overline{X}_{t-1} &\sim N(0.7 Z_{t-1} + 0.2 X_{t-1}, \, 1)\\
X_t \mid \overline{Z}_{t}, \overline{X}_{t-1} &\sim \operatorname{Bernoulli}(\operatorname{expit}(0.5 X_{t-1} + 0.25 Z_t))\\ 
P(Y_t = 1 \mid do(\overline{X}_{t}), W) &= \exp\{ - X_t + W/10\},
\end{align*}
and assume that there is a Gaussian pair-copula linking $Y_t$ with $Z_{t-1},Z_t$,
which also depends upon $W$.  We assume that the parameter for the $Y_t$-$Z_{t-1}$
interaction is 0.2, and that for $Y_t$-$Z_{t}$ given $Z_{t-1}$ is 0.3.

```{r formulas}
forms <- list(W ~ 1,
              Z ~ Z_l1 + X_l1,
              X ~ X_l1 + Z_l0,
              Y ~ W + X_l0,
              list(Y = list(Z ~ W)))
fams <- list(3, 1, 5, 3, c(1))
pars <- list(W = list(beta=0, phi=1/2),
             Z = list(beta=c(0,0.7,0.2), phi=1),
             X = list(beta=c(-0.5,0.25,0.5)), 
             Y = list(beta=c(1,-1/10,1/5), phi=1),
             cop = list(Y=list(Z=list(beta=c(0.5, 0.0))))
)
link <- list("log", "identity", "logit", "inverse")
```
suppressWarnings(library(survey))
suppressWarnings(library(survival))
suppressWarnings(library(ipw))
gamma1 <- 0.1
gamma2 <- 0.15
gamma3 <- 0.1
gamma4 <- 0.3
gamma5 <- 0.3


rho_to_beta <- function(rho){
  x <- (rho + 1)/2
  return(log(x/(1-x)))
}


n <- 1e5
qtls <- data.frame(X1 = runif(n), X2 = runif(n), B1 = runif(n), B2 = runif(n),
                   Z1_0 = runif(n), A_0 = runif(n))
qtls[["Z2|Z1_0"]] = runif(n)

dat <- data.frame(X1=qnorm(qtls$X1,mean = 0, sd = 1),
                  X2 = qbinom(qtls$X2, size = 1, prob = 0.5))
dat$B1 <- qnorm(qtls$B1, 0.4*dat$X2 -0.2, sd = 1)
dat$B2 <- qnorm(qtls$B2, 0.2*dat$X1, sd = 1)

dat[["Z1_0"]] <- qnorm(qtls[["Z1_0"]], 0.2*dat$X1, sd = 1)
dat[["Z2_0"]] <- qbinom(qtls[["Z2|Z1_0"]], 1, expit(-0.2 + 0.4 * dat$X2))
dat[["A_0"]] = qbinom(qtls[["A_0"]], 1, expit(-1 + 0.1 * dat$X1 + 0.15*dat$X2 + 0.1*dat$B1 + 0.3*dat[["Z1_0"]] + 0.3* dat[["Z2_0"]]))



formulas <- list(list(X1 ~ 1, X2 ~ 1,B1 ~ X2, B2 ~ X1),
                 list(Z1 ~  B2 + Z1_l1 + A_l1, Z2 ~ B2 + Z2_l1 + A_l1),
                 A ~ X1 + X2 + B1 + Z1_l0 + Z2_l0 + A_l1,
                 Y ~ A_l0,
                 list(copZ1 ~ 1, copZ2 ~ 1))

family <- list(list(1,5,1,1),
               list(1,5),
               5,
               3,
               list(2,2))
pars <- list(X1 = list(beta = 0, phi = 1),
             X2 = list(beta = 0), # expit(0) = 0.5
             B1 = list(beta = c(-0.2,0.4),phi = 1),
             B2 = list(beta = c(0, 0.2), phi = 1),
             
             Z1 = list(beta=c(0.3,0.4, 0.7,-0.6 ), phi=1),
             Z2 = list(beta = c(-0.2,0.4,1,-0.6)),
             A = list(beta=c(-1, gamma1, gamma2, gamma3, gamma4, gamma5, 1)),
             Y = list(beta = c(0.5,0.2), phi = 1), # negative 1/exp
             copZ1 = list(beta= rho_to_beta(-0.6), par2 = 5),
             copZ2 = list(beta = rho_to_beta(0.2), par2 = 5))

dat2 <- msm_samp(T = 10,dat = dat, qtls = qtls,
                 formulas = formulas,
                 family = family,
                 pars = pars,
                 link = list(list("identity", "logit", "identity", "identity"),
                             list("identity", "logit"),
                             "logit", "log"))

# dat <- as.data.table(dat)
datl <- surv_to_long(dat2)


temp <- ipwtm(
  exposure = A,
  family = "binomial",
  numerator = ~ 1,
  denominator = ~ Z1 +Z2,
  link = "logit",
  id = id,
  timevar = t,
  type = "all",
  data = datl)


datl$wt <- temp$ipw.weights
glmY <- suppressWarnings(coxph(Surv(t, t_stop,Y)~ A, id = id, data = datl,
                               weights = datl$wt, timefix = FALSE))
sumY <- summary(glmY)




